Static Discharge of Higher-Order Interposition

### 2013-10-01 pldi2014/paper.tex
Lots of changes.

\newtheorem{thm}{Theorem}

@@

@@

%\begin{abstract}
%This is the text of the abstract.
%\end{abstract}

@@

The implementation of familiar language features is sometimes accomplished through alien constructs.
Higher-order contracts (cite Felleisen and Findler) are now familiar constructs of languages but their implementation in Racket, which must cope with other language features--perhaps incidentally, is in terms of value wrappers \emph{impersonators} and \emph{chaperones}.
Nevertheless, analysis tools have access to more of certain kinds of information after compilation and so have the potential to prove more and different properties about programs.
Specifically, analysis tools for Racket have access to the whole program after module compilation through a \emph{demodularization} process.
In Racket, data contracts are implemented at the bytecode level using interposition primitives--that is, values are wrapped and these wrappers intercept arguments and results.
Because functions are data in a higher-order language, special care has been given to enforcing safe and sound function contracts.
Of course, codomain contract compliance cannot be decided.
In consequence, codomain contract checks are deferred until the function actually delivers a value.

Racket program lifecycle--the tipping of various types of information

This goal can be sought at each of the various stages of the Racket program lifecycle.
Initially, programs are organized into modules, and dynamic behavior can be gleaned from contract declarations at module boundaries.\emph{cite Tobin-Hochstadt and Van Horn}
Each module is compiled in isolation to bytecode; the bytecode compiler performs optimizations which rely on at most basic flow analysis are performed (constant-folding, 
``The bytecode compiler applies all standard optimizations, such as constant propagation, constant folding, inlining, and dead-code elimination.''
The journey to this stage discards high-level information and doesn't acquire much new information, so analysis here is probably fruitless.
Racket programs are organized into modules which consist of a sequence of import declarations, export declarations, definitions, and expressions.
When a module A is required in the program, the module is \emph{visited} which consists evaluating definition bodies and expressions to initialize exported variables.
Because Racket has rich facilities for syntactic extension, execution of code must be carefully controlled, especially since modules can be compiled separately.
(cite You Want It When)
Racket modules are compiled to bytecode. This

implementation of contracts within racket using interposition

talk about blame and the method Racket uses to assign blame--the compilation process, etc.

We discuss impersonators and chaperones in more depth in Section 2. We introduce the syntax and semantics of chapcalc, a core calculus with facilities for impersonating and chaperoning closures, in Section 3. We present two theorems which establish the conditions for the sound promotion or discharge of a selected chaperone creation site in Section 4. We present a static analysis which can judge when these conditions hold in Section 5. We present performance benchmarks of programs subject to this static analysis in Section 6. We discuss how this work fits with higher-order contracts and contract discharge in Section 7. We discuss limitations and future directions for this work in Section 8.

@@

\subsection{Impersonators and Chaperones}

Interposition is accomplished via two primitives, \emph{impersonators} and \emph{chaperones}.

Impersonators intercept arguments and results and can modify them arbitrarily before delivering them to their original destination.
Because such an ability amounts to mutation, impersonators are restricted to values where mutation can be observed, including functions.

Chaperones refine impersonators in that their interposition is restricted to installing chaperones on arguments and performing side effects--in particular, raising exceptions to signal contract violations and assign blame.

Higher-order contract checking can introduce enormous overhead to program execution--in real-world programs, as high as 75\% \emph{?}--but we expect that many of the contracts will never be violated.
The static detection and discharge of these contracts would therefore increase program performance significantly without compromising safety.

chaperone calculus, including the specific semantics of impersonators and chaperones

We consider a small language with facilities for impersonating and chaperoning procedures, chapcalc.
We inherit the constraints of impersonators and chaperones from Racket.
To accommodate these constraints, chapcalc \scheme{let}- and \scheme{letrec}-binding forms have the same structure as \scheme{λ}-binding forms.
For example,
\begin{schemedisplay}
(let ([(a) (f x)])
  ...)
\end{schemedisplay}
binds the single result of \scheme{(f x)} to \scheme{a}, and \scheme{raise}s if \scheme{(f x)} does not result in exactly one value (just as an application \scheme{raise}s if its arity is incompatible the number of arguments supplied).
In contrast,
\begin{schemedisplay}
(let ([a (f x)])
  ...)
\end{schemedisplay}
binds a list of any number of results of \scheme{(f x)} to \scheme{a}.
Finally,
\begin{schemedisplay}
(let ([(a . b) (f x)])
  ...)
\end{schemedisplay}
binds the first result of \scheme{(f x)} to \scheme{a} and a list of all remaining results to \scheme{b}.

This is the grammar for chapcalc0.
\begin{schemedisplay}
(grammar
  [e ((e e ...)
      (app-values e e)
      (values e ...)
      (chaperone-operator e e e)
      (impersonate-operator e e e)
      (let ([s e]) e)
      (letrec ([s e]) e)
      (λ s e)
      (if e e e)
      (or e e)
      (and e e)
      (raise)
      x
      v
      p)]
  [p (= < + - * integer? boolean? pair? null? cons car cdr null)]
  [v (n b)]
  [n (0 1 -1 2 -2 ...)]
  [b (#t #f)]
  [s (x () (x . s))])
\end{schemedisplay}

Expressions \scheme{e} have the following forms.
Procedure applications of the form \scheme{(e e ...)} apply a procedure to a static sequence of arguments.
The \scheme{app-values} form applies a procedure to a list of values as if the constitent values were supplied statically.
The \scheme{values} forms returns any number of values as results.

The \scheme{or} form is strict in its first expression position.
The \scheme{and} form is strict in its first expression position.

To make the handling of wrapper procedures easier, call and return sites are symmetric in chapcalc. The binding specification of a $\lambda$ form (statically) designates a finite number of arguments to bind individually and an optional \emph{rest} parameter which binds any remaining arguments, as can that of a \scheme{let} or \scheme{letrec} form. If the binding specification of a form does not have a \scheme{rest} parameter, its arity is the number of arguments it binds--a natural number.

When speaking about operators, we will refer to its arguments arity, which is the set of acceptable numbers of arguments, and its results arity, which is the set of possible numbers of results.

The \scheme{chaperone-operator} operator takes as arguments: an operator \scheme{f} to be chaperoned, an operator \scheme{neg-wrap} which must have the same argument arities as \scheme{f} and produce as many results as arguments given, and an operator \scheme{pos-wrap} which must have the same result arities as \scheme{f} and accept as many arguments as results produced.

For each wrapper, then, there are two properties which must be enforced.
Of these four properties, only one--that \scheme{neg-wrap} has the same argument arity as \scheme{f}--can be enforced at chaperone creation (and Racket takes this approach). For simplicity and understandability, we will initially defer enforcement of each of these properties until chaperone invocation.

Definition: call an impersonator value \emph{well-behaved} with respect to a program if it never \scheme{raise}s as a result of an arity mismatch and its wrappers preserve arities.

[m] -> neg-wrap -> [m] -> f -> [n] -> pos-wrap -> [n]

Definition: call a chaperone value well-behaved with respect to a program if it is a well-behaved impersonator and every result of its wrappers is \scheme{chaperone-of?} the corresponding argument.

% chaperone-of? within the language can change the behavior, right? if it was a chaperone before, removing the chaperone makes it still chaperone-of? it's chaperone? that makes it possible to detect changes, but that is discouraged, and we will leave it out of our language.

[m] -> neg-wrap -> C[m] -> f -> [n] -> pos-wrap -> C[n]

Chaperone values are branded with the label of their creation site. In any given program, there are a finite number of these creation sites and labels. In fact, it is these creation sites that we are concerned with. The output of an analysis is a map from labels to transformations (no-op, promotion, discharge). Once we apply these transformations, we recover optimization opportunities as data dependency decreases.

We call a wrapper function $g$ an \emph{originator} of a chaperone $c_0$ if $c_0$ is created in within the extent of $g$. Chaperone values are also branded with their parent chaperon

Definition: call a creation site good0 if all of the chaperone values which flow from it are well-behaved.

If the wrapper returns an additional result, its first result is a function (that may depend on input) that has argument and result arities of the result arities of the wrapped function.

\begin{thm}[Selective Chaperone Discharge Theorem]
A good and nice chaperone installment site can soundly be replaced by its first argument (where ``soundly'' means ``with no change to the meaning of the program'').
\end{thm}

\begin{proof}
The only way that a chaperone can change the meaning of a program is by causing an error, raising an error, or causing the creation of a chaperone with the same property.

By definition, all chaperones which flow from a good and nice chaperone installment site are good and nice meaning that they do not cause an error, raise an error, or cause the creation of a chaperone with the same property.
\end{proof}

@@

Because the result wrapper is produced at each invocation, it may depend on

Strickland presents a chaperone erasure theorem which states that if a program with chaperones evaluates to a value, the same program with chaperones erased--that is, with all chaperone installment sites replaced with the first argument--will evaluate to the same value. The purpose of this theorem seems to be to make a statement about the nature of chaperones--that, in a well-behaved program, they do not change meaning. Our theorems give the conditions of when a given chaperone installment site does not change the meaning of the program.

This language is sufficient to express two theorems.

Definition: Call an impersonator \emph{well-behaved} with respect to a program if its application never \scheme{raise}s.
(only talking about arity and special runtime stuff here)

Well-behavedness is defined with respect to a program--as opposed to assigning the property in isolation--because an impersonator may misbehave in situations never encountered within a particular program.

In general, we cannot decide whether a particular impersonator--in context or isolation--has this property.

Definition: Call a chaperone wrapping function \emph{well-behaved} with respect to a program if it only possibly installs chaperones on its arguments.

Definition: Call a chaperone \emph{well-behaved} with respect to a program if its application never \scheme{raise}s.

Definition: Call a chaperone \emph{satisfied} with respect to a program if it never \emph{raise}s and

Def: call a chaperone good0 wrt a program if it at most chaperones its arguments.
Def: call a chaperone-installment site good0 wrt a program if all chaperones which flow from it are good0.
Def: call a chaperone good1 wrt a program if it is good0 and its installing chaperone, if any, is good1.
Def: call a chaperone-installment site good1 wrt a program if it is good0 and the chaperone-installment sites of its installing chaperones, if any, are good1.
Thm: a good1 chaperone-installment site can be promoted to an impersonator-installment site if its installing chaperone site, if any, is promoted.

@@

Def: call a chaperone good2 wrt a program if its application never raises a runtime (arity) error, its wrapper never raises.
Def: call a chaperone good3 wrt a program if it is good2 and the chaperones it installed, if any, are good3.
Thm: if all chaperones which flow from a chaperone-installment site are good3, that site can be replaced with its first argument (the until-now-wrapped value).

\begin{schemedisplay}
(grammar
  [e ((e e ...)
      (app-values e e)
      (chaperone-operator e e)
      (impersonate-operator e e)
      (let ([s e]) e)
      (letrec ([s e]) e)
      (if e e e)
      (or e e)
      (and e e)
      (raise e)
      (handle ([x e]) e)
      (λ s e)
      x)]
  [p (= < + - * integer? boolean?)]
  [v (n b)]
  [n (0 1 -1 2 -2 ...)]
  [b (#t #f)]
  [s (x () (x . s))])
\end{schemedisplay}

@@

>>>>

\mathchardef\mhyphen="2D
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}

@@

\newcommand{\chapcalc}[0]{\ensuremath{\lambda_{\mathcal{C}}}}

@@

\begin{abstract}
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
This is the text of the abstract.
\end{abstract}

@@

% what is the problem?
Higher-order behavioral contracts impose time and space costs on programs.

It takes time to determine compliance of first-order values as the first-order contract can be an arbitrary predicate.
Additionally, functions may ascribe contracts to their higher-order arguments which themselves do similarly, even recursively.
Execution time of the program proper can be dwarfed by that of the enforcement of its specification \cite{strickland2012chaperones}.
%There isn't much a calculus can do to solve the time issue, especially in the face of dependent contracts.

Because contract compliance of higher-order values cannot be determined in general at ascription time, compliance checks must be attached to values to be performed later.
These attachments may accumulate, even changing the asymptotic space complexity of the program [cite herman].

This space issue is well-known and a few different approaches have been taken to deal with it:
Herman's coercion calculus, Wadler and Sieks' threesome calculus, the space-efficient contracts author
These do well to soundly preserve the asymptotic space complexity of the program and perhaps the time as well.

Much of the reason that these contracts cannot be statically discharged is that programs are written and compiled modularly and the greatest amount of knowledge a module can have about its imports \emph{is} their behavioral contract.\footnote{Indeed, it can be less than this, as when the behavioral specification is not codified entirely within the program.} Recognizing that contracts constitute total formal knowledge, Tobin-Hochstadt and Van Horn \cite{tobin2012higher} design an analysis to reason about this knowledge.
%TB and VH do this at the highest language level. This analysis is designed to account for exactly the shortcomings of a modular program: a dearth of information.

%Ideally, we would like to statically discharge the contract.
%[Determining statically whether a function meets a type is the purpose of a type system. static analysis is perceived to be more heavyweight and can prove more about the program.]
%This approach is well-known.

Whole-program compilers (cite MLTon) are able to perform thorough inter-procedural and inter-modular analysis and effect significant program optimizations.

However, static analysis generally performs best when given access to the entire program.
We can provide this access if we are willing to analyze Racket bytecode (which, contrary to its name, is a fairly rich language).
At the bytecode level, programs can be \emph{demodularized} into the conceptual single module which a static analyzer expects.
Higher-order contracts are manifest in Racket bytecode by \emph{interposition}.

Recent advances in static analysis [cite Earl] % blech
allow conservative analysis of the stack.
The motivation for this was to provide an abstract garbage collector a set of live frames, disregarding any stack structure latent in the control flow graph.
The pushdown characterization immediately admits a finite description of an unbounded stack as a regular expression.
This description alone is enough to allow us to reason statically about features based on stack inspection, such as security checks [cite] and dynamic binding [cite].
We will use it to approximate blame on a misbehaving program.
After considering all possible sources of such blame, a straightforward analysis allows us to alter contract sites, shifting some of the dynamic burden to the static.

% thus, in order to soundly discharge contracts from the original program, we must soundly discharge impersonators and chaperones.
%The implementation of familiar language features is sometimes accomplished through alien constructs.
% (Because the feature must be situated particularly within the already existing features of the language.)
%Contracts are a common feature of dynamically-typed languages.
%Higher-order contracts \cite{findler2002contracts} increase the domain of contracts to higher-order values.
%Of course, codomain compliance of a higher-order contract cannot be decided at attachment time.
%Instead, contracts are decomposed and positioned to intercept the arguments and results of application--the earliest possible time that contract compliance can be determined.
%This \emph{de facto} wrapping of deferred contract checks imposes space and time overhead.
%Wadler and his cronies and Herman and his cronies \cite{wadler2009well,herman2010space} show how the accumulation of wrappers can grow unbounded and offer various solutions to ameliorate this.
%Determining contract compliance can make up a significant portion of reduction.

%Higher-order contracts \cite{findler2002contracts} are now familiar constructs of languages but their implementation in Racket, which must cope with other language features--perhaps incidentally, is in terms of value wrappers.
% \emph{impersonators} and \emph{chaperones}.
%In Racket, data contracts are implemented at the bytecode level using interposition primitives--that is, values are wrapped and these wrappers intercept arguments and results.
%Because functions are data in a higher-order language, special care has been given to enforcing safe and sound function contracts.
%Of course, codomain contract compliance cannot be decided.
%In consequence, codomain contract checks are deferred until the function actually delivers a value.

%Nevertheless, analysis tools have access to more of certain kinds of information after compilation and so have the potential to prove more and different properties about programs.
%Specifically, analysis tools for Racket have access to the whole program after module compilation through a \emph{demodularization} process.

%Racket program lifecycle--the tipping of various types of information

% a static analysis defines a class of program transformations
% if the analysis is sound, the transformation is meaning-preserving
% are these two preceding statements true?

%This goal can be sought at each of the various stages of the Racket program lifecycle.
%Initially, programs are organized into modules, and dynamic behavior can be gleaned from contract declarations at module boundaries.\cite{tobin2012higher}
%Each module is compiled in isolation to bytecode; the bytecode compiler performs optimizations which rely on at most basic flow analysis are performed (constant-folding, 
%``The bytecode compiler applies all standard optimizations, such as constant propagation, constant folding, inlining, and dead-code elimination.''
%The journey to this stage discards high-level information and doesn't acquire much new information, so analysis here is probably fruitless.
%Racket programs are organized into modules which consist of a sequence of import declarations, export declarations, definitions, and expressions.
%When a module A is required in the program, the module is \emph{visited} which consists evaluating definition bodies and expressions to initialize exported variables.
%Because Racket has rich facilities for syntactic extension, execution of code must be carefully controlled, especially since modules can be compiled separately.
%(cite You Want It When)
%Racket modules are compiled to bytecode. This

%implementation of contracts within racket using interposition

%talk about blame and the method Racket uses to assign blame--the compilation process, etc.

We discuss interposition in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for impersonating and chaperoning closures, in Section 3.
We present two theorems which establish the conditions for the sound promotion or discharge of a selected chaperone creation site in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We present performance benchmarks of programs subject to this static analysis in Section 6.
We discuss how this work fits with higher-order contracts and contract discharge in Section 7.
We conclude by discussing the limitations and future directions of this work in Section 8.

@@

The \scheme{chaperone-procedure} procedure takes a procedure \scheme{p} to be chaperoned as its first argument and a procedure \scheme{w} to wrap \scheme{p} as its second argument.
The argument and result arities of \scheme{p} and \scheme{w} must be compatible.
This means that the argument arity of \scheme{w} must be the same as
%\footnote{Racket procedures can include both mandatory and optional keyword arguments. When considering these, it is necessary only that the set of mandatory keyword arguments of \scheme{p} include those of \scheme{w} and the set of all keyword arguments of \scheme{p} be included in.}
The result of a \scheme{chaperone-procedure} call is a procedure which, upon invocation, will deliver the arguments to \scheme{w}.
\begin{enumerate}
\item The $n$ arguments are delivered to \scheme{w}.
\item If \scheme{w} returns $n$ results, the results are 
\end{enumerate}

\section{Interposition}

Behavioral contracts in Racket are manifest at run time by \emph{impersonators} and \emph{chaperones}.

An impersonator wraps a value and intercepts operations on them, potentially redirecting them.
In the case of procedures--first-class values in Racket--impersonators intercept arguments and can alter or replace them before passing them to the procedure they wrap.
As they pass these arguments, they have the opportunity to register for interest in intercepting the results of the application by including an extra argument: a callback procedure which will be applied to the results of the inner procedure.

In many ways, impersonators are unconstrained in their behavior: they can perform side-effects, even aborting the pending application; they can alter arguments arbitrarily before passing them along, a capability tantamount to mutation.
However, with respect to procedures, their behavior must meet a minimum standard of predictability: wrapper procedures must return the same number of values as given (except for the optional extra argument mentioned earlier); the arity of a wrapper procedure must include the arity of the procedure to be wrapped, a condition enforced at impersonator installation time.

Chaperones refine impersonators by trading behavioral freedom for reasonability.
Well-behaved procedure chaperones cannot alter intercepted arguments arbitrarily; they must return them as given or possibly with chaperones installed on them.
Chaperones share the ability of impersonators to perform side-effects.
In particular, they retain the ability to raise exceptions, the mechanism by which contract violations are signalled.

Chaperones are intended to be treated like the original value.
% box?, procedure?, etc. will respond the same way
% they evaluate equal? to the values they wrap, but not eq?
Procedure chaperones are considered to have the same arity of the procedures they wrap.
The \scheme{chaperone?} predicate is designated as the single way to detect chaperones at runtime, though its use is discouraged.

%the bytecode level in terms of interposition installation.
%Higher-order and mutable values are wrapped with \emph{impersonators} which redirect operations on the value and mediate breaches of expectations between the value and its context.
%Impersonators can redirect behavior arbitrarily, which may violate some expectations or invariants the user would like to establish, so contracts are expressed in terms of a restricted form of impersonators, \emph{chaperones}.
%Chaperones are limited to chaperoning values that pass through their membrane and performing side effects (particularly the announcement of a contract violation in the form of an exception).

%Run-time interposition effects a run-time cost: impersonators have to be well-behaved in certain ways the limitations of chaperones must by enforced dynamically.

%Soundly discharging a contract, then, amounts to preventing the installment of the chaperone in the first place.
%If that is not possible,

%Behavioral contracts on higher-order and mutable objects must be checked on every operation to which the contract pertains--an up-front check is not enough.
%Racket uses run-time interposition to implement value contracts in terms of two proxy primitives: \emph{impersonators} and \emph{chaperones}.

%An impersonator wraps a value and redirects operations on it.

%Impersonators wrap values and redirect operations in which they interpose.

%The implementation of value contracts in Racket is accomplished via \emph{impersonators} and \emph{chaperones}.
%Impersonators and chaperones wrap values and intercept

%The implementation of higher-order contracts in Racket is accomplished via procedure \emph{impersonators} and \emph{chaperones}.

%Procedure impersonators intercept application arguments and results and forward them to the procedure or context, respectively, although they may modify or replace them before doing so.
%Additionally, they may exhibit side effects.

%Function chaperones refine impersonators in that they are restricted in the modification they can apply to intercepted values.
%Procedure chaperones must return their arguments or chaperones of their arguments.

%Impersonators and chaperones thus constitute an interposed wrapper between values and their contexts, and so we refer to them as \emph{interposition}.

%Interposition is accomplished via two primitives, \emph{impersonators} and \emph{chaperones}.

%Impersonators intercept arguments and results and can modify them arbitrarily before delivering them to their original destination.
%Because such an ability amounts to mutation, impersonators are restricted to values where mutation can be observed, including functions.

%Chaperones refine impersonators in that their interposition is restricted to installing chaperones on arguments and performing side effects--in particular, raising exceptions to signal contract violations and assign blame.

%Higher-order contract checking can introduce enormous overhead to program execution--in real-world programs, as high as 75\% \emph{?}--but we expect that many of the contracts will never be violated.
%The static detection and discharge of these contracts would therefore increase program performance significantly without compromising safety.

%chaperone calculus, including the specific semantics of impersonators and chaperones

%the motivation to soundly discharge or promote a chaperone

%the conditions for a discharge to be sound

%a fixed-point type of algorithm to determine which can be erased
%erase the ones at the "bottom"
%run the analysis again
%keep going until there are none at the bottom
%must handle recursive sites, if that's possible
%we don't have to run the analysis again, just analyze the results

%the result of the analysis is a program transformation on labelled sites

%\section{Chaperone Calculus}

% Racket is big and we don't/can't analyze all of it, nor do we want to report on it, so we deal with a small calculus.
% Part of the purpose is to limit the language; part is to limit the exposition.
% We will clear what is safe and easy to add.

%We consider a small language with facilities for impersonating and chaperoning procedures, \chapcalc.

%We inherit the constraints of impersonators and chaperones from Racket.
%To accommodate these constraints, \chapcalc \scheme{let}- and \scheme{letrec}-binding forms have the same structure as \scheme{λ}-binding forms.
%For example,
%\begin{schemedisplay}
%(let ([(a) (f x)])
%  ...)
%\end{schemedisplay}
%binds the single result of \scheme{(f x)} to \scheme{a}, and \scheme{raise}s if \scheme{(f x)} does not result in exactly one value (just as an application \scheme{raise}s if its arity is incompatible the number of arguments supplied).
%In contrast,
%\begin{schemedisplay}
%(let ([a (f x)])
%  ...)
%\end{schemedisplay}
%binds a list of any number of results of \scheme{(f x)} to \scheme{a}.
%Finally,
%\begin{schemedisplay}
%(let ([(a . b) (f x)])
%  ...)
%\end{schemedisplay}
%binds the first result of \scheme{(f x)} to \scheme{a} and a list of all remaining %results to \scheme{b}.

%This is the grammar for \chapcalc 0.
%\begin{figure*}
%\input{calc0}
%\end{figure*}

%programs either evaluate to a value, evaluate to an error, or diverge.
%our theorem only deals with chaperones which don't affect whether the program evaluates to a value?
%a program could diverge on chaperone creation (evaluating the guards) or chaperone application (applying the guards).
%this is undecidable in general.
%it is possible to remove a chaperone and have some programs that errored out evaluate to a value (because the guards may artificially restrict the domain or codomain of the function).
%it is possible to remove a chaperone and have some programs that errored out diverge (because the error came before divergence).
%if a program evaluates to a value then no chaperone errored out and no chaperone diverged.
%we also are only changing chaperones that don't error in any circumstances.
%therefore, if a program evaluates to a value with a given input before erasure or promotion, it will evaluate to that value after erasure or promotion.
%if a program evaluates to an error with a given input, it will evaluate to an error after, since we will not erase or promote a chaperone that /could/ throw an error, for a sound analysis.
%however, programs that would have diverged otherwise may evaluate to a value or an error.

%Figure whatever defines the grammar for \chapcalc 0 expressions. Expressions \scheme{e} have the following forms
%Procedure applications of the form \scheme{(e e ...)} apply a procedure to a static sequence of arguments.
%The \scheme{app-values} form applies a procedure to a list of values as if the constitent values were supplied statically.
%The \scheme{values} forms returns any number of values as results.

%The \scheme{or} form is strict in its first expression position.
%The \scheme{and} form is strict in its first expression position.

%To make the handling of wrapper procedures easier, call and return sites are symmetric in \chapcalc. The binding specification of a $\lambda$ form (statically) designates a finite number of arguments to bind individually and an optional \emph{rest} parameter which binds any remaining arguments, as can that of a \scheme{let} or \scheme{letrec} form. If the binding specification of a form does not have a \scheme{rest} parameter, its arity is the number of arguments it binds--a natural number.

%When speaking about operators, we will refer to its arguments arity, which is the set of acceptable numbers of arguments, and its results arity, which is the set of possible numbers of results.

%Figure whatever2 defines the evaluation contexts for \chapcalc 0 expressions.
%Evaluation contexts are either a hole which is the destination for the evaluation of the current redex...

%Figure whatever3 defines the reduction rules for \chapcalc 0.
%Rule blank applies when the redex is an application form and the operator is a closure.
%The result is blank.
%Rule blank applies when the redex is an application form and the operator is a primitive.
%The result is blank or an error.
%Rule blank applies when the redex is an application form and the operator is a chaperone.
%If the arguments arity of the negative guard does not include the number of arguments given, an error is raised.
%The context is extended with a frame annotated with the value and ``negative'' and the arguments given.
%When the results return, etc.

%Decomposition of a term into context and redex is unique and the rule conditions are disjoint, so these rules define a function on terms.

%Let $\longrightarrow\longrightarrow$ be the reflexive, transitive closure of this relation.
%This defines an evaluation function.
%Programs in \chapcalc 0 can result in a value, an error, or divergence.
%\begin{definition}[Soundness]
%If a program $p$ with input $x$ evaluates to a value $v$, then $T[p]$ with input $x$ evaluates to $v$.
%If a program $p$ with input $x$ results in an error, then $T[p]$ with input $x$ results in an error.
%\end{definition}

%While blame is an apt description of the judgment we are making, our notion of blame is not the same as [Findler and Wadler].
%The category of behaviors for which we must account include whether errors arise as a result of chaperone creation or chaperone application.
%In each case, we are concerned with the creation site of the chaperone application.
%If an error arises as a result

%While there is only one mechanism to provoke errors, the erroneous result of the program resides in one of four classes depending on the state of execution at the time of provocation.
%(Bad format below.)
%\begin{enumerate}
%\item If evaluation of a \scheme{chaperone-operator} application is occurring, the error is transformed to a blame assignment of that site.
%\item If evaluation of the negative guard of a chaperone is occurring, the error becomes \scheme{(blame -L)}.
%\item If evaluation of the positive guard of a chaperone is occurring, the error becomes ...
%\item none means none
%\item if the negative guard does more than chaperone arguments, blame should be assigned to the guard in some sense, indicating that not only can the guard not be discharged, but the chaperone site also cannot be demoted to an impersonator. (remember the types of errors we need to track!)
%\end{enumerate}
%An error is transformed according to the innermost part of the context which meets the criteria.

% make a redex model of the language
% generate random program
% run analysis on random program
% transform according to analysis
% ensure that the result is the same
% generate random program that takes input (either lists, functions, or numbers)
% run analysis
% transform according to analysis
% ensure that the result is the same for a set of random inputs

%Our transformation preserves valid evaluation and preserves erroneous evaluation.
%We do not (and cannot) say anything about divergent evaluation.
%$T[p]$ may converge on some input $x$ where $p$ diverged.

%By this definition, the identity transformation is sound. We will take the position that if the creation or application of a chaperone value \emph{may} result in an error, we will not discharge it. Only when our conservative analysis indicates that an error cannot be raised will we discharge it.

%As an error is raised, it starts to unwind the stack. If a negative frame is encountered, blame is allocated to over p (negative p). Similarly for a positive frame. If neither type of frame is encountered, the result is simply \scheme{error}.

%The \scheme{chaperone-operator} operator takes as arguments: an operator \scheme{f} to be chaperoned, an operator \scheme{neg-wrap} which must have the same argument arities as \scheme{f} and produce as many results as arguments given, and an operator \scheme{pos-wrap} which must have the same result arities as \scheme{f} and accept as many arguments as results produced.

%For each wrapper, then, there are two properties which must be enforced.
%Of these four properties, only one--that \scheme{neg-wrap} has the same argument arity as \scheme{f}--can be enforced at chaperone creation (and Racket takes this approach). For simplicity and understandability, we will initially defer enforcement of each of these properties until chaperone invocation.

%Definition: call an impersonator value \emph{well-behaved} with respect to a program if it never \scheme{raise}s as a result of an arity mismatch and its wrappers preserve arities.

%[m] -> neg-wrap -> [m] -> f -> [n] -> pos-wrap -> [n]

%talk about the run-time cost of chaperones and the optimizations we would like to perform:

%STILL NEED CHAPERONE VALUES TO TRACK THEIR CREATING CHAPERONE (OR AT LEAST ITS LABEL).

%Guards can fail by raising an error. (This error is translated into a blame error by the context?)
%Guards can also fail by changing their arguments more than chaperoning.

%A guard that never raises an error and does not chaperone any of its arguments is ``OK''.
%A chaperone with an ``OK'' negative guard and an ``OK'' positive guard is ``OK''.
%A guard that never raises an error and only creates ``OK'' chaperones is ``OK''.
%An ``OK'' guard can be replaced with \scheme{values}.
%A chaperone creation site that never raises and with only ``OK'' guards is ``OK''.
%An ``OK'' chaperone creation site can be replaced with its first argument.

%If a program $p$ with input $x$ evaluates to a value $v$, then program $p$ with an ``OK'' chaperone creation site replaced with its first argument evaluates to $v$.

%Definition: call a chaperone value well-behaved with respect to a program if it is a well-behaved impersonator and every result of its wrappers is \scheme{chaperone-of?} the corresponding argument.

%\subsection{Handling \emph{handle}}

%Exception handlers prevent certain errors from escaping a particular dynamic extent.
%Consider the program
%\begin{schemedisplay}
%(handle
%  (chaperone-operator + (raise) (raise))
%  -)
%\end{schemedisplay}
%wherein an error is always raised and the handler always triggered.
%Even though an exception never escapes the dynamic extent of the \scheme{handle} expression, replacing the \scheme{chaperone-operator} application with its first argument results in a different program. 
%This means that we must consider whether a \scheme{chaperone-operator} application ever raises an error, and not simply if that error escapes a delimited extent.
% being able to catch errors means that errors can change the meaning of a program arbitrarily--even if it results in a value.
% this means that we need to preserve errors
% chaperone-of? within the language can change the behavior, right? if it was a chaperone before, removing the chaperone makes it still chaperone-of? it's chaperone? that makes it possible to detect changes, but that is discouraged, and we will leave it out of our language.

%[m] -> neg-wrap -> C[m] -> f -> [n] -> pos-wrap -> C[n]

%Chaperone values are branded with the label of their creation site. In any given program, there are a finite number of these creation sites and labels. In fact, it is these creation sites that we are concerned with. The output of an analysis is a map from labels to transformations (no-op, promotion, discharge). Once we apply these transformations, we recover optimization opportunities as data dependency decreases.

%We call a wrapper function $g$ an \emph{originator} of a chaperone $c_0$ if $c_0$ is created in within the extent of $g$. Chaperone values are also branded with their parent chaperon

%Definition: call a creation site good0 if all of the chaperone values which flow from it are well-behaved.

%If the wrapper returns an additional result, its first result is a function (that may depend on input) that has argument and result arities of the result arities of the wrapped function.

%\begin{theorem}[Selective Chaperone Discharge Theorem]
%A good and nice chaperone installment site can soundly be replaced by its first argument %(where ``soundly'' means ``with no change to the meaning of the program'').
%\end{theorem}

%\begin{proof}
%The only way that a chaperone can change the meaning of a program is by causing an error, raising an error, or causing the creation of a chaperone with the same property.

%By definition, all chaperones which flow from a good and nice chaperone installment site are good and nice meaning that they do not cause an error, raise an error, or cause the creation of a chaperone with the same property.
%\end{proof}

@@

%Because the result wrapper is produced at each invocation, it may depend on

%Strickland presents a chaperone erasure theorem which states that if a program with chaperones evaluates to a value, the same program with chaperones erased--that is, with all chaperone installment sites replaced with the first argument--will evaluate to the same value. The purpose of this theorem seems to be to make a statement about the nature of chaperones--that, in a well-behaved program, they do not change meaning. Our theorems give the conditions of when a given chaperone installment site does not change the meaning of the program.

%This language is sufficient to express two theorems.

%Definition: Call an impersonator \emph{well-behaved} with respect to a program if its application never \scheme{raise}s.
%(only talking about arity and special runtime stuff here)

%Well-behavedness is defined with respect to a program--as opposed to assigning the property in isolation--because an impersonator may misbehave in situations never encountered within a particular program.

%In general, we cannot decide whether a particular impersonator--in context or isolation--has this property.

%Definition: Call a chaperone wrapping function \emph{well-behaved} with respect to a program if it only possibly installs chaperones on its arguments.

%Definition: Call a chaperone \emph{well-behaved} with respect to a program if its application never \scheme{raise}s.

%Definition: Call a chaperone \emph{satisfied} with respect to a program if it never \emph{raise}s and

%Def: call a chaperone good0 wrt a program if it at most chaperones its arguments.
%Def: call a chaperone-installment site good0 wrt a program if all chaperones which flow from it are good0.
%Def: call a chaperone good1 wrt a program if it is good0 and its installing chaperone, if any, is good1.
%Def: call a chaperone-installment site good1 wrt a program if it is good0 and the chaperone-installment sites of its installing chaperones, if any, are good1.
%Thm: a good1 chaperone-installment site can be promoted to an impersonator-installment site if its installing chaperone site, if any, is promoted.

@@

%Def: call a chaperone good2 wrt a program if its application never raises a runtime (arity) error, its wrapper never raises.
%Def: call a chaperone good3 wrt a program if it is good2 and the chaperones it installed, if any, are good3.
%Thm: if all chaperones which flow from a chaperone-installment site are good3, that site can be replaced with its first argument (the until-now-wrapped value).

@@

% the concrete interpreter correctly assigns blame but there is a problem:
% if the chaperone-operator is wrapped in a function, then it can only be changed if /all/ the uses of that function meet the requirements. it doesn't matter if the store is going to be garbage-collected--all the values are branded with its syntactic label.
% one thing we could do is attempt to distinguish calls into classes and divert all calls of a certain class to a different function. for instance, if a chaperone value as it goes around the program is well-behaved, can we look where it is created and change it in isolation somehow?
